Theories of deliberated judgment

### 2022-07-22 simple.tex
hbadness for biblio
Deliberated choice

On the other hand, since at least the seminal work of Fishkin, another sort of judgment involving subjective desirability is emerging as also worth studying. These are judgments given by subjects after arguments in favor and against different possible stance have been considered. Such judgments may differ from immediate judgments of desirability given in natural conditions. Indeed, individuals may change opinion on the desirability of some course of action, and therefore revise their choice, while learning about the properties of objects, the consequences of some acts (or empirical knowledge useful to estimate the likelihood of some consequences), or the logical or empirical impossibilities between consequences (for example, in some situations, you can’t both pay less taxes and have better government provided services). This is particularly likely to happen when choosing among non everyday objects or acts whose consequences are not well known by the subject.

@@

To define deliberated judgments, one needs to start with the set of arguments under consideration, and the way the subject gets to know them, called hereafter the “exposure protocol”. As a result, we need always talk about the deliberated judgment \emph{given some exposure protocol}. The notion of argument invoked here is an extremely large one, including texts, images, sounds, experiences (a hiking trip), and so on: anything that can be transmitted to a subject qualifies as a possible argument. Such an “argument” need not even have the logical structure of what is considered a proper argument in argumentation theory, which permits to avoid having to take position on what is a correct argument, and permits to study wide enough conceptions of deliberated judgment, including those involving no paternalism: anything that influences the judgment of a subject may a priori be considered worth including in at least some conceptions of deliberated judgment.

The dependence of the deliberated judgment concept on the exposure protocol yields a first qualification about the concept of deliberated judgments defined here when one desires to use it normatively: it is only as reasonable to call the resulting judgments “deliberate” in a normative sense as the exposure protocol is normatively relevant to deliberation. A sufficiently restricted, or even purposefully chosen, set of arguments may yield judgments that are only deliberate in the technical sense used here, but that pertain more to brain-washing than to reflective judgments. While it is indeed a long-term goal of this research stream to yield a normatively relevant conception of preference (appropriate, for example, for recommendation), it is not a claim of this article that anything that corresponds to deliberated judgments as defined here qualify as being appropriate for recommendation (or for any other normative use): much depends on the chosen exposure protocol. The goal of this article is, more modestly, one of a conceptual and practical clarification. It proposes a sharp distinction between the empirical content and the normative content of the concept of deliberated judgments: the definition proposed here transfers the full normative weight on the choice of the exposure protocol; and leaves the rest of the disagreements to empirical settling. This is methodologically akin, mutatis mutandis, to the axiomatic study of justice such as done in social choice, only replacing mathematical deduction with empirical claims. In social choice, the choice of axioms bears the full normative weight; once axioms are chosen, it is a matter of deduction which social choice rules (if any) correspond to that choice. In our case, once an exposure protocol is chosen, it is a pure empirical matter which deliberated judgments result.
In supplement to a possibly helpful conceptual clarification, such a sharp separation may reveal practically useful for separation of concern as it permits to study deliberated judgments empirically under various exposure protocols independently of possible persistent disagreements about which norms are more appropriate for deliberation in which circumstances.

@@

An exposure protocol is a tuple $(I, \allargs, {\allleadsto})$, where $I$ is a set of individuals, $\allargs$ is a set of arguments, and ${\allleadsto}$ is the behavior of the individuals as a function of the arguments they have been exposed to.

$\allargs$ is a set of arguments. 
Throughout the article, $\N$ includes zero. %and $\N^* = \N \setminus \set{0}$. 
Given $j, l \in \N$, the notation $\intvl{j, l}$ represents $\set{k \in \N \suchthat j ≤ k ≤ l}$ (thus $\forall j \in \N: \intvl{j, 0} = \emptyset$).
Define $\allhist = \bigcup_{k \in \N} \allargs^{\intvl{1, k}}$ as the finite sequences of arguments. A sequence of arguments is called an argumentative path, or simply a path.
A generic element of $\allhist$ is denoted by $\hist$ and called an argumentative path, or simply a path when this raises no ambiguity. Its length is denoted by $\card{\hist} \in \N$.
Thus, given an argumentative path $\hist \in \allhist$, $\forall k \in \intvl{1, \card{\hist}}$, $\hist_k \in \allargs$, and $\hist = \emptyset ⇔ \card{\hist} = 0$.

The function ${\allleadsto}: I × \allhist → \powersetz{\Phi}$ associates to each individual and path a non-empty subset of $\Phi$. The case $\Phi$ is used to represent indifferent choice, in case this is observable.
The observation that $\phi$ is considered among the best by $i$ considering $\alpha$ is represented by $\alpha \ileadsto \phi$, considering ${\ileadsto} \subseteq \allhist × \Phi$ as a binary relation.
As usual with binary relations, $\hist \nileadsto \prop$ means $(\hist, \prop) \notin {\ileadsto}$.

% which associates to each $i \in I$ an element of $\Phi$ or nothing, with the semantics that this decision is the right one for that individual or that the theory abstains from recommending.

@@

\bibliography{simple}

>>>>

On the other hand, since at least the seminal work of Fishkin, another sort of judgment involving subjective desirability is emerging as also worth studying. These are judgments given by subjects after arguments in favor and against different possible stance have been considered. Such judgments may differ from immediate judgments of desirability given in natural conditions. Indeed, individuals may change opinion on the desirability of some course of action, and therefore revise their choice, while learning about the properties of objects, the consequences of some acts (or empirical knowledge useful to estimate the likelihood of some consequences), or the logical or empirical impossibilities between consequences. This is particularly likely to happen when choosing among non everyday objects or acts whose consequences are not well known by the subject.

@@

To define deliberated judgments, one needs to start with the set of arguments under consideration, and the way the subject gets to know them, called hereafter the “exposure protocol”. As a result, we need always talk about the deliberated judgment \emph{given some exposure protocol}. The notion of argument invoked here is an extremely large one, including texts, images, sounds, experiences (a hiking trip), and so on: anything that can be transmitted to a subject qualifies as a possible argument. Such an “argument” need not even have the logical structure of what is considered a proper argument in argumentation theory, which permits to avoid having to take position on what is a correct argument, and permits to study wide enough conceptions of deliberated judgment, including those involving no paternalism: anything that influences the judgment of a subject may a priori be considered worth including in at least some conceptions of deliberated judgment, thus including whatever some could consider “bad reasoning” or “bad reasons”.

The dependence of the deliberated judgment concept on the exposure protocol yields a first qualification about the concept of deliberated judgments defined here when one desires to use it normatively: it is only as reasonable to call the resulting judgments “deliberate” in a normative sense as the exposure protocol is normatively relevant to deliberation. A sufficiently restricted, or even purposefully chosen, set of arguments may yield judgments that are only deliberate in the technical sense used here, but that pertain more to brain-washing than to reflective judgments. While it is indeed a long-term goal of this research program to yield a normatively relevant conception of preference (appropriate, for example, for recommendation), it is not a claim of this article that anything that corresponds to deliberated judgments as defined here qualify as being appropriate for recommendation (or for any other normative use): much depends on the chosen exposure protocol. The goal of this article is, more modestly, one of a conceptual and practical clarification. It proposes a sharp distinction between the empirical content and the normative content of the concept of deliberated judgments: the definition proposed here transfers the full normative weight on the choice of the exposure protocol; and leaves the rest of the disagreements to empirical settling. This is methodologically akin, mutatis mutandis, to the axiomatic study of justice such as done in social choice, only replacing mathematical deduction with empirical claims. In social choice, the choice of axioms bears the full normative weight; once axioms are chosen, it is a matter of deduction which social choice rules (if any) correspond to that choice. In our case, once an exposure protocol is chosen, it is a pure empirical matter which deliberated judgments result.
In supplement to a possibly helpful conceptual clarification, such a sharp separation may reveal practically useful for separation of concern as it permits to study deliberated judgments empirically under various exposure protocols independently of possible persistent disagreements about which norms are more appropriate for deliberation in which circumstances.

@@

\subsection{Exposure protocol}
An exposure protocol is a tuple $(I, \allargs, {\allleadsto})$, where $I$ is a set of individuals, $\allargs$ is a set of arguments, and ${\allleadsto}$ is the behavior of the individuals as a function of the arguments they have been exposed to.

While $I$ and $\allargs$ are fundamental sets whose elements are considered elementary, a few notations and concepts are required to define ${\allleadsto}$.
Throughout the article, $\N$ includes zero. %and $\N^* = \N \setminus \set{0}$. 
Given $j, l \in \N$, the notation $\intvl{j, l}$ represents $\set{k \in \N \suchthat j ≤ k ≤ l}$ (thus $\forall j \in \N: \intvl{j, 0} = \emptyset$).
Define $\allhist = \bigcup_{k \in \N} \allargs^{\intvl{1, k}}$ as the finite sequences of arguments, including the empty sequence. 
A generic element of $\allhist$ (a sequence of arguments) is denoted by $\hist$ and called an argumentative path, or simply a path when this raises no ambiguity. Its length is denoted by $\card{\hist} \in \N$.
Thus, given an argumentative path $\hist \in \allhist$, $\forall k \in \intvl{1, \card{\hist}}$, $\hist_k \in \allargs$, and $\hist = \emptyset ⇔ \card{\hist} = 0$.

The function ${\allleadsto}: I × \allhist → \powersetz{\Phi}$ associates to each individual and path a non-empty subset of $\Phi$, ${\allleadsto}(i, \alpha)$ representing the choice of $i$ following exposure to the arguemnts of $\alpha$, in order. The case ${\allleadsto}(i, \alpha) = \Phi$ is used to represent indifferent choice, in case this is considered observable.
If $\phi \in {\allleadsto}(i, \alpha)$, we say that $\phi$ is considered as one of the best options by $i$ considering $\alpha$. 
It is also denoted by $\alpha \ileadsto \phi$, considering ${\ileadsto} \subseteq \allhist × \Phi$ as a binary relation.
As usual with binary relations, $\hist \nileadsto \prop$ means $(\hist, \prop) \notin {\ileadsto}$ or equivalently $\phi \notin {\allleadsto}(i, \alpha)$.

\subsection{Deliberated choice}
We can now define the deliberated choice of $i$ given the exposure protocol $(I, \allargs, {\allleadsto})$.
It is denoted $\Phi_i$ and is a subset of $\Phi$, with the semantics that $\phi \in \Phi_i$ means that $\phi$ is considered by $i$, after deliberation, one of the best options, given the exposure protocol. In other words, if $\Phi_i = \Phi$, the individual considers both $\phi$ and $¬\phi$ equally worth after deliberation and if $\Phi_i = \set{\phi}$, the individual considers $\phi$ the strictly best option among the proposed set, after deliberation (and similarly for $\Phi_i = \set{¬\phi}$).

@@

\hbadness=10000
\bibliography{simple}
